---
title: "mtcars mpg prediction"
author: "Ning Zhang"
date: "December 20, 2015"
output: html_document
---

You work for Motor Trend, a magazine about the automobile industry. Looking at a data set of a collection of cars, they are interested in exploring the relationship between a set of variables and miles per gallon (MPG) (outcome). They are particularly interested in the following two questions:

“Is an automatic or manual transmission better for MPG”
"Quantify the MPG difference between automatic and manual transmissions"

```{r}
data(mtcars)
str(mtcars)
```

```{r}
#calculate correlations between features, and get rid of those highly related.
correlations <- cor(mtcars[,-1])
correlations <- correlations - diag(1, 10, 10)
"""
cyl <-> disp, hp, wt, vs,
disp <-> hp, drat, wt,vs,
hp <-> qsec, vs, carb,
drat <-> wt, am,
qsec <-> vs,
cyl, disp, hp, wt, vs, drat, qsec, carb, am,
"""

```

```{r}


fit_general <- lm(mpg ~ . - disp, data=mtcars)
summary(fit_general)

```

```{r}
fit_am <- lm(mpg ~ am, data=mtcars)
summary(fit_am)
fit_am_carb <- lm(mpg ~ am + carb, data = mtcars)
summary(fit_am_carb)
fit_am_carb_hp <- lm(mpg ~ am + carb + hp, data = mtcars)
summary(fit_am_carb_hp)
fit_am_hp <- lm(mpg ~ am + hp, data = mtcars)
summary(fit_am_hp)
plot(fit_am_hp)


fit_am_hp_wt <- lm(mpg ~ am + hp + wt, data=mtcars)
summary(fit_am_hp_wt)

fit_am_hp_qsec <- lm(mpg ~ am + hp + qsec, data=mtcars)
summary(fit_am_hp_qsec)

mtcars$hpSq <- mtcars$hp^2
fit_am_hpSq <- lm(mpg ~ am + hpSq, data=mtcars)
summary(fit_am_hpSq)

```

```{r}
data(mtcars)
fit <- lm(mpg ~ ., data=mtcars)
summary(fit)
fit <- update(fit, . ~ . - cyl)
summary(fit)
fit <- update(fit, . ~ . - vs)
summary(fit)

fit <- update(fit, . ~ . - carb)
summary(fit)
fit <- update(fit, . ~ . - gear)
summary(fit)
fit <- update(fit, . ~ . - drat)
summary(fit)

fit <- update(fit, . ~ . - disp)
summary(fit)
fit <- update(fit, . ~ . - hp)
summary(fit)
plot(fit)



mtcars$qsecRoot <- sqrt(mtcars$qsec)
fit_final <- lm(mpg ~ wt + qsec + am + qsecRoot, data=mtcars)
summary(fit_final)

fit <- lm(mpg ~ wt + am + qsec - 1, data = mtcars)
summary(fit)

```

```{r}
plot(predict(fit), mtcars$mpg)
line(predict(fit), mtcars$mpg)
abline(0.5207, 0.9476)
```

problems to solve:

How to look at the fit plot?

